---
title: FC Barcelona Enters a New Era
author: Nolan Hemmett
date: '2022-04-29'
slug: fc-barcelona-enters-a-new-era
categories: []
tags: []
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(worldfootballR)
library(cowplot)
library(scales)
library(ggbump)
library(waffle)
```



```{r, message=FALSE, warning=FALSE, echo=FALSE}
# function to extract La Liga match results data
LaLiga_2022 <- get_match_results(country = "ESP", gender = "M", season_end_year = 2022, tier = "1st") 
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
Barcapool<- fb_big5_advanced_season_stats(season_end_year = (2021:2022), stat_type = "standard", team_or_player = "player") %>% filter(Comp== "La Liga") %>% filter(Squad== "Barcelona")
```




```{r, message=FALSE, warning=FALSE, echo=FALSE}

Barcagoals_21<- Barcapool %>% 
  filter(Season_End_Year == "2021") %>% 
  select(Season_End_Year, Player, Gls) %>% 
  arrange(desc(Gls)) %>% 
  top_n(5)

Messi <- Barcagoals_21 %>% filter(Player == "Lionel Messi")
Griezmann <- Barcagoals_21 %>% filter(Player == "Antoine Griezmann")
Dembele <- Barcagoals_21 %>% filter(Player == "Ousmane Dembélé")
Fati <- Barcagoals_21 %>% filter(Player == "Ansu Fati")
Alba <- Barcagoals_21 %>% filter(Player == "Jordi Alba")
deJong <- Barcagoals_21 %>% filter(Player == "Frenkie de Jong")
Pedri <- Barcagoals_21 %>% filter(Player == "Pedri")
Trincao <- Barcagoals_21 %>% filter(Player == "Francisco Trincão")

bar1 <- ggplot() + 
  geom_bar(data=Barcagoals_21, aes(x=reorder(Player, Gls), weight=Gls)) + 
  geom_bar(data=Messi, aes(x=reorder(Player, Gls), weight=Gls), fill="#A50044") +
  geom_bar(data=Griezmann, aes(x=reorder(Player, Gls), weight=Gls), fill="#A50044") +
  geom_bar(data=Dembele, aes(x=reorder(Player, Gls), weight=Gls), fill="#A50044") +
  geom_bar(data=Fati, aes(x=reorder(Player, Gls), weight=Gls), fill="#A50044") +
  geom_bar(data=Alba, aes(x=reorder(Player, Gls), weight=Gls), fill="#A50044") +
  geom_bar(data=deJong, aes(x=reorder(Player, Gls), weight=Gls), fill="#A50044") +
  geom_bar(data=Pedri, aes(x=reorder(Player, Gls), weight=Gls), fill="#A50044") +
  geom_bar(data=Trincao, aes(x=reorder(Player, Gls), weight=Gls), fill="#A50044") +
  coord_flip() + 
  labs(title="A New Age For Barcelona", subtitle = "With a new manager in Xavi, Barcelona must fill a gaping hole after Messi left in 2021.",  x="2021 Goal Scorers", y="Goals", caption = "") + 
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10), 
    panel.grid.minor = element_blank()
    )
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
Barcagoals_22<- Barcapool %>% 
  filter(Season_End_Year == "2022") %>% 
  select(Season_End_Year, Player, Gls) %>% 
  arrange(desc(Gls)) %>% 
  top_n(5)

Depay <- Barcagoals_22 %>% filter(Player == "Memphis Depay")
Aubameyang <- Barcagoals_22 %>% filter(Player == "Pierre-Emerick Aubameyang")
Luuk <- Barcagoals_22 %>% filter(Player == "Luuk de Jong")
Araujo <- Barcagoals_22 %>% filter(Player == "Ronald Araújo")
Torres <- Barcagoals_22 %>% filter(Player == "Ferrán Torres")

bar2 <- ggplot() + 
   geom_bar(data=Barcagoals_22, aes(x=reorder(Player, Gls), weight=Gls)) + 
  geom_bar(data=Depay, aes(x=reorder(Player, Gls), weight=Gls), fill="#004D98") +
  geom_bar(data=Aubameyang, aes(x=reorder(Player, Gls), weight=Gls), fill="#004D98") +
  geom_bar(data=Luuk, aes(x=reorder(Player, Gls), weight=Gls), fill="#004D98") +
  geom_bar(data=Araujo, aes(x=reorder(Player, Gls), weight=Gls), fill="#004D98") +
  geom_bar(data=Torres, aes(x=reorder(Player, Gls), weight=Gls), fill="#004D98") +
  scale_y_continuous(breaks = pretty_breaks()) + 
  coord_flip() + 
  labs(title="", subtitle = "",  x="2022 Goal Scorers", y="Goals", caption = "By: Nolan Hemmett | Source: Football Reference") + 
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10), 
    panel.grid.minor = element_blank()
    )
```

As far as crazy seasons go, FC Barcelona in La Liga have had a rollercoaster of a 2021/22 season. It all started with rumors of Lionel Messi being linked with a move to Paris Saint-Germain F.C in France. 

Messi sprouted in Barcelona’s club academy and has spent his entire career in Barcelona, so no one saw it coming when the rumor was true and just like that the backbone of Barcelona for 21 years was gone. 

All eyes were on Barcelona as they entered a new era, the year after Messi.

Antoine Griezmann also moved back to Atlético Madrid, widening the void that manager Ronald Koeman was desperate to fill. 


```{r, message=FALSE, warning=FALSE, echo=FALSE}
plot_grid(bar1,bar2)
```


```{r, message=FALSE, warning=FALSE, echo=FALSE}
table21 <- read_csv("table21.csv")
```

Barcelona was bleeding after Messi left, the club found themselves falling down the table with no hope in sight. Barcelona fell to 9th after a humiliating loss to Rayo Vallecano. After which, then current manager and ex Barcelona player, Ronald Koeman was fired. 

This left the door open for another ex-player and Barcelona legend, Xavi, to bring the club back to its former glory. All without Lionel Messi. 

Xavi’s instant impact is nothing short of a miracle. Xavi implemented stricter rules, and brought back the “Barca way”, playing smooth and fast. 

After Xavi was given the manager job, Barcelona would remain undefeated for three months. They were able to climb their way all the way back to second on the table. 

The height of Xavi’s new Barcelona was in the 249th El Clasico, a match between super rivals: Barcelona and Real Madrid. Real Madrid were the favorites to win. But a brace from Pierre Emerick-Aubameyang, a goal from Ferran Torres, and another from Ronald Araújo made the game a lopsided 4-0 to Barcelona. No one saw that coming.


```{r, message=FALSE, warning=FALSE, echo=FALSE}
laligarank <- table21 %>% 
  filter(matchday == max(matchday)) %>%
  top_n(10, pts)

teams_wk33 <- laligarank$squad

top_10_33 <- table21 %>% 
  filter(squad %in% teams_wk33)

barcelona <- table21 %>% 
  filter(squad == 'Barcelona')

ggplot() +
  geom_bump(data = top_10_33, aes(x = matchday, y = rk, group = squad), colour = 'grey') +
  geom_bump(data = barcelona, aes(x = matchday, y = rk, group = squad), colour = '#A50044') +
   geom_point(data = top_10_33, aes(x = matchday, y = rk, group = squad), size = 1, colour = 'grey') +
   geom_point(data = barcelona, aes(x = matchday, y = rk), size = 1, colour = '#A50044') +
   geom_text(data = barcelona %>% filter(matchday == max(matchday)), aes(x = matchday - .2, y=rk, label = squad), size = 3, hjust = -.2) +
  geom_vline(xintercept = 13)+
  annotate('text', x = 16.75, y = 17, label = 'Xavi hired as manager.', size=3)+
  scale_y_reverse()+
  scale_x_continuous(limits = c(1,35), breaks = c(0, 5, 10, 15, 20, 25, 30)) +
  labs(
    title="Barcelona is Back To Winning Form", 
    subtitle="New manager, Xavi, has brought Barcelona back to second place after disasterous start to season.", 
    caption = "By: Nolan Hemmett  |  Source: Football Reference",
    y= "Rank",
    x = "Matchday") +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank()
    )+
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size = 10),
    plot.caption = element_text(size = 8),
    panel.grid.minor = element_blank()
    )
```



```{r, message=FALSE, warning=FALSE, echo=FALSE}
BarcaGD<- table21 %>% filter(squad == "Barcelona")
MadridGD<- table21 %>% filter(squad == "Real Madrid")
```


The height of Xavi’s new Barcelona was in the 249th El Clasico, a match between super rivals: Barcelona and Real Madrid. Real Madrid were the favorites to win. 

A brace from Pierre Emerick-Aubameyang, a goal from Ferran Torres, and another from Ronald Araújo made the game a lopsided 4-0 to Barcelona. No one saw that coming.

```{r, message=FALSE, warning=FALSE, echo=FALSE}

ggplot() +
  geom_line(data = table21, aes(x = matchday, y = g_diff, group = squad), color = "grey") +
  geom_line(data = BarcaGD, aes(x = matchday, y = g_diff), color = "#A50044") +
   annotate('text', x = 26.5, y = 25, label = 'Barcelona', size=3)+
  geom_line(data = MadridGD, aes(x = matchday, y = g_diff), color = "#FEBE10") + 
  annotate('text', x = 14.5, y = 25, label = 'Real Madrid', size=3)+
  scale_y_continuous(limits = c(-30, 50), breaks = c(-30,-20,-10,0,10,20,30,40))+
  labs(y = "Goal Differential", x = "Matchday",
       title = "Barcelona Is Closing The Gap",
       subtitle = "After losing at home to Rayo Vallecano, Barcelona are currently 15 points behind their rival on top of the table,\nReal Madrid. The progress is clear but it won't be enough to finish on top. ",
       caption = "By: Nolan Hemmett  |  Source: Football Reference"
       ) +
  theme_minimal() + 
  theme(
  panel.grid.major = element_blank()
  )+
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 8),
    panel.grid.minor = element_blank()
  )
  
```
Barcelona was eliminated from the Champions League and next year have to fight their way back in the Europa League. They will be back.

With a great mind like Xavi and great signings in Aubameyang and young talent like Gerrard Pique and Gavi, Barcelona are in great hands. Now they look forward into their future hungry to prove their worth in the post-Messi era. 